---
title: "More more on Normal Distributions"
author: "your name"
date: "2024-10-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Math 2265 Sec 4.1 Normal Distribution Part III

- Work as a group!
- You will need to replace `"ans"` or `your_answer` in the source code
- Update your name in L3
- Add your group members' name below; students may lose one point if Question 0 is unanswered
- Make sure you save and `knit` your work (to html or pdf) before submitting it to Canvas

---

### Goal

- Learn the function that computes the percentile from given x or its Z-score.
- Learn the function that computes x or its Z-score or from given percentile.
- Master the functions `pnorm` and `qnorm`.

---

### Question 0. Who are your group members? (List their first names should be sufficient)

**Answer:** 

  1. `<name_1>`
  1. `<name_2>`

### Load Packages

```{r, echo=FALSE, include=FALSE}
library(openintro)
library(ggplot2)
if (!tinytex::is_tinytex()) {
  tinytex::install_tinytex()
} # this ensures the necessary package is installed for knit to pdf
library(tinytex)  # to knit to PDF
library(mosaic)
set.seed(2265)
```

---

### If you need more time to get used to `Markdown`, use the `Visual` mode.

The icon is located in the upper-left corner next to `source`. 

---

### Example 1: Visualizing percentiles in the normal distribution with `xpnorm`

(a) P(Z < -1.1)

```{r}
xpnorm(-1.1, mean = 0, sd = 1)
```

(b) P(Z < 0.35)

```{r}
xpnorm(0.35, mean = 0, sd = 1)
```

(c) P(Z > 0.5)

```{r}
xpnorm(0.5, mean = 0, sd = 1)
```

Caveat: `xpnorm` computes both P(Z<0.5) and P(Z>0.5).

(d) P(|Z| < 2) = P (-2 < Z < 2)

```{r}
xpnorm(c(-2,2), mean = 0, sd = 1)
```

Caveat: This does not compute P(|Z|< 2), but P(Z<-2) and P(Z<2) for -2, 2. So, either we take the difference or use

```{r}
pnorm(2, mean=0,sd=1) - pnorm(-2,mean=0,sd=1)
```

### Example 2: Visualizing percentiles in normal distributions with `xpnorm`

(a) x = 98, $\mu = 92.6$, $\sigma = 3.6$

```{r}
xpnorm(98, mean=92.6, sd=3.6)
```
Note that the Z-score, 1.5, is computed. 

(b) x = 89, $\mu = 92.6$, $\sigma = 3.6$

```{r}
xpnorm(89, mean=92.6, sd=3.6)
```

### Example 3: Visualizing Z-scores from Percentiles

What is the Z-score for the 80th percentile with mean = 0 and standard deviation = 1?

```{r}
qnorm(0.8, mean = 0, sd = 1)
```

Now check out the output of `xqnorm`.

```{r}
xqnorm(0.8, mean = 0, sd = 1)
```

---

## HW Problem

### Speeding on the I-5 Freeway

The distribution of passenger vehicle speeds traveling on the Interstate 5 Freeway (I-5) in California is nearly normal with:

- Mean speed ($\mu$): 72.6 miles/hour
- Standard deviation ($\sigma$): 4.78 miles/hour

### (a) What percent of passenger vehicles travel slower than 80 miles/hour?

First, compute the Z-score for $x = 80$:

```{r}
z_80 <- (80 - 72.6) / 4.78
z_80
```

Now, compute the percentile using the standard normal distribution:

```{r}
p_80 <- pnorm(z_80, mean = 0, sd = 1)
p_80
```

Alternatively, you can compute the percentile directly without calculating the Z-score:

```{r}
p_80_direct <- pnorm(80, mean = 72.6, sd = 4.78)
p_80_direct
```

### (b) What percent of passenger vehicles travel between 60 and 80 miles/hour?

First, compute the Z-score for $x = 80$:

```{r}
z_60 <- (80 - 72.6) / 4.78
z_60
```

Now, compute the percentile using the standard normal distribution:

```{r}
p_60 <- pnorm(z_60, mean = 0, sd = 1)
p_60
```

Alternatively, compute directly without Z-scores:

```{r}
p_60_direct <- pnorm(60, mean = 72.6, sd = 4.78)
p_60_direct
```

Compute the percentile (we already computed `p_80_direct` above).

```{r}
p_80_direct - p_60_direct
```

### (c) How fast do the fastest 5% of passenger vehicles travel?

We need to find the speed corresponding to the 95th percentile (since the fastest 5% are above this speed).

Compute the speed using the inverse normal distribution function `qnorm`:

```{r}
speed_95th <- qnorm(0.95, mean = 72.6, sd = 4.78)
speed_95th
```

### (d) Approximate what percentage of the passenger vehicles travel above the speed limit of 70 miles/hour.

Compute the Z-score for $x = 70$:

```{r}
z_70 <- (70 - 72.6) / 4.78
z_70
```

Compute the percentile up to 70 miles/hour:

```{r}
p_70 <- pnorm(z_70, mean = 0, sd = 1)
p_70
```

Alternatively, compute directly:

```{r}
p_70_direct <- pnorm(70, mean = 72.6, sd = 4.78)
p_70_direct
```

Compute the percentage of vehicles traveling above 70 miles/hour:

```{r}
percentile_above_70 <- 1 - p_70
```

Answers:

| Question                                   | Answer              |
|--------------------------------------------|---------------------|
| (a) Percent traveling slower than 80 mph  | 93.94%              |
| (b) Percent traveling between 60 and 80 mph | 93.48%              |
| (c) Speed of the fastest 5% of vehicles   | Above 80.46 mph     |
| (d) Percent traveling above the speed limit of 70 mph | 70.68%      |

Here is visualization with `xpnorm` and `xqnorm`:

```{r}
xpnorm(80, mean = 72.6, sd = 4.78)
xpnorm(c(60,80), mean = 72.6, sd = 4.78)
xqnorm(0.95, mean = 72.6, sd = 4.78)
xpnorm(70, mean = 72.6, sd = 4.78)
```

### Task 6. Knit your code and check your outcomes.

You are only allowed to upload pdf or html

### Task 7. Check your answer 

```{r}
# Correct answers for comparison
correct_answers <- list(
  a = 0.9394,  # Percent traveling slower than 80 mph
  b = 0.9348,  # Percent traveling between 60 and 80 mph
  c = 80.46,   # Speed of the fastest 5% of vehicles
  d = 0.7068   # Percent traveling above the speed limit of 70 mph
)

# Part (a): Check percentile for slower than 80 mph
if (abs(p_80_direct - correct_answers$a) < 0.0005) {
  print("Part (a): Your answer is correct")
} else {
  print("Part (a): Check your answer")
}

# Part (b): Check percentile for between 60 and 80 mph
if (abs(p_80_direct - p_60_direct - correct_answers$b) < 0.0005) {
  print("Part (b): Your answer is correct")
} else {
  print("Part (b): Check your answer")
}

# Part (c): Check speed of the fastest 5%
if (abs(speed_95th - correct_answers$c) < 0.05) {
  print("Part (c): Your answer is correct")
} else {
  print("Part (c): Check your answer")
}

# Part (d): Check percentage traveling above 70 mph
if (abs(percentile_above_70 - correct_answers$d) < 0.0005) {
  print("Part (d): Your answer is correct")
} else {
  print("Part (d): Check your answer")
}
```

### Share your work and help your group members before uploading your work to Canvas