---
title: "Random Variables"
author: "your name"
date: "2024-09-26"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Math 2265 Chatch-up

- Work as a group!
- You will need to replace `"ans"` or `your_answer` in the source code or answer questions
- Update your name in L3
- Make sure your save and `knit` your work (to html or pdf) before submitting it to Canvas

### Question 0. Who are your group members? (List their first names should be sufficient)

**Answer: ** 

  1. `<name_1>`
  1. `<name_2>`

### Load Packages

```{r, echo=FALSE}
library(openintro)
library(ggplot2)
if (!tinytex::is_tinytex()) {
  tinytex::install_tinytex()
} # this ensures the necessary package is installed for knit to pdf
library(tinytex)  # to knit to PDF
set.seed(2265)
```

---

### If you need more time to get used to `Markdown`, use the `Visual` mode.

The icon is located in the upper-left corner next to `source`. 

---

### Task 1

Go to the data set list and choose one you are most interested in:

https://www.openintro.org/data/

Use `str` and `head` to checkout the dataset.

```{r}
# Define the winnings for X and their respective probabilities P(X)
str(your_answer)
```

```{r}
# Define the winnings for X and their respective probabilities P(X)
head(your_answer)
```

---

### Task 2

Choose a numerical variable and plot its histogram.

```{r}
# Define the winnings for X and their respective probabilities P(X)
ggplot(data=your_answer, aes(x=your_answer)) +
  geom_histogram()
```


### Task 3

Choose a categorical varialbe and make a table.


```{r}
# Define the winnings for X and their respective probabilities P(X)
table(your_answer$your_answer)
```

---
### Task 4. Knit your code and check your outcomes.

You are only allowed to upload pdf or html

### Share your work and help your group members before uploading your work to Canvas
