---
title: "Linear Regression II"
author: "your name"
date: "2024-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(openintro)
library(ggplot2)
if (!tinytex::is_tinytex()) {
  latex_installed <- Sys.which("tlmgr") != ""
  
  if (!latex_installed) {
    tinytex::install_tinytex()
  } else {
    message("A LaTeX distribution already exists on the system.")
  }
} # this ensures TinyTeX is installed only if no LaTeX distribution is found
library(tinytex)  # to knit to PDF
set.seed(2265)
```

## Math 2265 Chapter 8. Linear Regression

- Work as a group!
- You will need to replace `"ans"` or `your_answer` in the source code
- Update your name in L3
- Add your group members' name below; students may lose one point if Question 0 is unanswered
- Make sure you save and `knit` your work (to html or pdf) before submitting it to Canvas
- Please only submit your work if you attended the class and worked with other students; this is not an online course

---

### Goal

1. Inference using a  linear model
1. Apply linear regression to a real work data set

---

### Question 0. Who are your group members? (List their first names)

**Answer:** 

  1. `<name_1>`
  1. `<name_2>`

---

### If you need more time to get used to `Markdown`, use the `Visual` mode.

The icon is located in the upper-left corner next to `source`. 

---

### `duke_forest`: Sale prices of houses in Duke Forest, Durham, NC

Data Set: `duke_forest`.

We will work with the `duke_forest` data set.
Write a script to check the number of observations and the names of the variables of the data set. 

```{r}
# write your code 

``` 

Recall we use linear regression for a pair of numerical variables, explanatory (horizontal, often denoted by x) and response variable (vertical, often denoted by y).   


### Task 

Save all the numerical variable names as an R vector to the R variable `numerical_variables` below 

- Just copy the names from the previous output and paste them as parameters of the vector 
- Strings (e.g., variable names) need to be enclosed with quotation marks "variable_name"

```{r}
numerical_variable <- c("variable_name1","variable_name2",...,...)
numerical_variable 
```

Among the variables we will use `price` as the response variable. 
The goal is to find the **most suitable explanatory variable** for linear regression.
The word *most* requires justification, and we will use two crietria: 

1. The correlation coefficient $R$. 
2. The sum of squared residuals with respect to the least squares line.

## Best with respect to the correlation coefficient

To do this, we will need to compute the correlation coefficients of `price` over all the rest of the numerical variables you found above. Use the following block to compute the correlation coefficients. 

```{r}
### 
cor_1 <- cor(duke_forest$price, your_answer)
cor_2 <- 
cor_3 <- 
cor_4 <-
cor_5 <-
cor_6 <-
cor_7 <-
cor_8 <-
# use as many as you need and delete unused ones

cor_vec <- c(cor_1, cor_2, cor_3, your_answer, your_answer, ... ) 
cor_vec
```

One of the results may be "NA" since that variable has missing value(s), and we won't use it for this worksheet.

Use `sort`, `max`, or `min` to determine the best one. Recall when determining the best, we need to consider their absolute values. 

```{r}
sort(abs(cor_vec))
```

---

**Question:** Which variables are linearly associated with the variable `price`? 

Write the top two:

1. "your_answer"
2. "your_answer"

---

## Best with respect to the least squares line

To do this, we will use `lm`. However, to use linear regression, we need to check that the variables are linearly associated with `price`. We can check it with their scatter plots. 

In the next cell, make scatter plots for each of the explanatory variables. 

```{r}
### Use as many as you need and delete unused ones
ggplot(data=duke_forest, aes(x=your_answer,y=price))+
  geom_point()

ggplot(data=duke_forest, aes(x=your_answer,y=price))+
  geom_point()

ggplot(data=duke_forest, aes(x=your_answer,y=price))+
  geom_point()

ggplot(data=duke_forest, aes(x=your_answer,y=price))+
  geom_point()

ggplot(data=duke_forest, aes(x=your_answer,y=price))+
  geom_point()
```

---

**Question:** Which two variables have the strongest linear association with `price`. 

Answer:

1. your_answer
2. your_answer

---

Now, we will use `lm` to find the least squares line and the sum of squared residuals for these two variables. 

### Task

Save the linear model for your 1st choice

```{r}
model1 <- lm( price ~ your_answer, data=duke_forest)
summary(model1) 
```

Compute the sum of the squared residuals and save it to `ssr1`.

```{r}
ssr1 <- sum( your_answer) 
ssr1
```

Save the linear model for your 2nd choice

```{r}
model2 <- lm( price ~ your_answer, data=duke_forest)
summary(model2) 
```

Compute the sum of the squared residuals and save it to `ssr2`.

```{r}
# ssr2 <- sum( your_answer) 
ssr2 <- sum( residuals(model2)^2 )
ssr2
```

Execute the next line to compare the results.

```{r}
if(ssr1 < ssr2) {
  print("Your 1st choice has lower error. Use this to make a prediction in the next part.")
} else {
  print("Your 2nd choice has lower error. Use this to make a prediction in the next part.")
}
```

---

### Question:

Which variable explains the responds variable `price` better?

Answer:

---

### Make prediction based on your model

First plot the least squares line. Use your answer above as your explanatory variable. 

```{r}
# ggplot(data=duke_forest, aes(x = your_answer, y = price)) +
#   geom_point() + 
#   geom_abline(intercept = your_answer, slope = your_answer, color='hotpink')

ggplot(data=duke_forest, aes(x = area, y = price)) +
  geom_point() + 
  geom_abline(intercept = your_answer , slope = your_answer, color='hotpink')
```

Make a prediction where your $x$ value is the median and mean of your variable. That is if your variable is called `my_expanatory_variable`, $x = median(my_expanatory_variable)$ or $x = mean(my_expanatory_variable)$.

```{r}
# x <- median(duke_forest$my_expanatory_variable)
# or use mean (or do both)
# x <- mean(duke_forest$my_expanatory_variable)
x <- median(duke_forest$area)

y_hat <- function(x) {
  # intercept + slope * x
  your_answer + your_answer * x
}

# You do not need to modify the following two lines. It will print the outcome.
print(paste("For", x, "y_hat(x) = ", y_hat(x)))
print(paste("That is, for a house of your variable ", x, "the estimated price is $", y_hat(x)))
```

## Please only submit your work if you attended the class and worked with other students; this is not an online course

## Complete the handout as a group, and turn it in to the instructor

### Share your work and help your group members before uploading your work to Canvas
