---
title: "Introduction to Normal Distribution"
author: "your name"
date: "2024-10-3"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Math 2265 Sec 4.1 Normal Distribution Part I

- Work as a group!
- You will need to replace `"ans"` or `your_answer` in the source code or answer questions
- Update your name in L3
- Add your group members name below
- Make sure your save and `knit` your work (to html or pdf) before submitting it to Canvas

### Question 0. Who are your group members? (List their first names should be sufficient)

**Answer: ** 

  1. `<name_1>`
  1. `<name_2>`

### Load Packages

```{r, echo=FALSE}
library(openintro)
library(ggplot2)
if (!tinytex::is_tinytex()) {
  tinytex::install_tinytex()
} # this ensures the necessary package is installed for knit to pdf
library(tinytex)  # to knit to PDF
library(mosaic)
set.seed(2265)
```

---

### If you need more time to get used to `Markdown`, use the `Visual` mode.

The icon is located in the upper-left corner next to `source`. 

---

### Example 1: Plot normal distributions

```{r}
curve(dnorm(x, mean = 0, sd = 1), from = -6, to = 6, 
      main = "Normal Distribution (PDF, mean = 0, sd = 1)",
      xlab = "x", ylab = "Density")
```
### Task 1: Plot some normal distributions with mean and standard deviation of your choice

You may need to change `from` and `to` parameters too. 

Example 1:
```{r}
curve(dnorm(x, mean = your_answer, sd = your_answer), from = -6, to = 6)
```

Example 2:

```{r}
curve(dnorm(x, mean = your_answer, sd = your_answer), from = -6, to = 6)
```

### Example 2: Normal distribution with colors

```{r}
# Plot the normal distribution curve
curve(dnorm(x, mean = 0, sd = 1), from = -4, to = 4, 
      main = "Normal Distribution with Shaded Area (x <= 0)",
      xlab = "x", ylab = "Density")

# Shade the area under the curve for x <= 0
x_vals <- seq(-4, 0, length.out = 100)  # Sequence of x values from -4 to 0
y_vals <- dnorm(x_vals, mean = 0, sd = 1)  # Corresponding y values for the curve
polygon(c(x_vals, 0), c(y_vals, 0), col = "lightblue", border = NA)  # Fill the area
```
This example is excellent but a bit involved. We will use another package `mosaic` for plotting normal distributions.

### Z-Score

$$
Z = \frac{x-\mu}{\rho},
$$
where $\mu$ is the mean and $\rho$ is the standard deviation.

### Task 2: Compute the Z-Score and percentile (with table)

[Add hw]

### Task 3: Re-do task 2 using the function `pnorm`


### Task 4: Test out the `xpnorm` funtion to plot normal distributions

---

### Task 5. Knit your code and check your outcomes.

You are only allowed to upload pdf or html

### Share your work and help your group members before uploading your work to Canvas
