---
title: "More on Normal Distributions"
author: "your name"
date: "2024-10-8"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Math 2265 Sec 4.1 Normal Distribution Part II

- Work as a group!
- You will need to replace `"ans"` or `your_answer` in the source code or answer questions
- Update your name in L3
- Add your group members name below; students may lose one point if Q0 is unanswered 
- Make sure your save and `knit` your work (to html or pdf) before submitting it to Canvas

---

### Goal

- Learn the function that computes the percentile from given x or its Z-score.
- Learn the function that computes x or its Z-score or from given percentile.

**Remark:**
- These processes are inverses to each other and understanding graphs (regions) is beneficial. 
- The Z-score is a way to transform any normal distribution to the standard normal distribution $\mu = 0$ and $\sigma = 1$.

---

### Question 0. Who are your group members? (List their first names should be sufficient)

**Answer: ** 

  1. `<name_1>`
  1. `<name_2>`

### Load Packages

```{r, echo=FALSE}
library(openintro)
#library(ggplot2)
if (!tinytex::is_tinytex()) {
  tinytex::install_tinytex()
} # this ensures the necessary package is installed for knit to pdf
library(tinytex)  # to knit to PDF
library(mosaic)
set.seed(2265)
```

---

### If you need more time to get used to `Markdown`, use the `Visual` mode.

The icon is located in the upper-left corner next to `source`. 

---

### Example 1: Percentage from Z-scores

When the Z-score is zero, then it splits the graph equally into two regions.

```{r}
pnorm(0, mean = 0, sd = 1)
```

In terms of the formula, 
$$
P(Z < 0) = 0.5
$$
### Task 1: Percentages from Z-scores

(a) z = -1.1

```{r}
pnorm(your_answer, mean = 0, sd = 1)
```

In terms of the formula, 
$$
P(Z yourAnswer ) = yourAnswer
$$
(b) z = 0.35

```{r}
pnorm(your_answer, mean = 0, sd = 1)
```
In terms of the formula, 
$$
P(Z yourAnswer) = yourAnswer
$$

(c) Do you remember how to compute $P(Z > 0.5)$?

```{r}
1 - pnorm(your_answer, mean = 0, sd = 1)
```
In terms of the formula, 
$$
P(Z yourAnswer) = yourAnswer
$$

### Example 2: Z-scores from Percentages

What is the Z-score for the 80th percentile with mean = 0 and standard deviation = 1?

We need to find the Z-score z such that $P(Z < z) = 0.8$, where $0.8 = 80/100$. 
We can use the `qnorm` function. The syntax is the same as `pnorm` but the first 
argument (parameter) is the percentile instead of the Z-score.

```{r}
qnorm(0.8, mean = 0, sd = 1)
```
That is (with marginal error), 
$$
P(Z < 0.8416212) = 0.8
$$
We can check it with `pnorm`.
```{r}
pnorm(0.8416212, mean = 0, sd = 1)
```

### Task 2:

(a) Fine the Z-score for 13.57 percentile.

```{r}
qnorm(your_answer, mean = 0, sd = 1)
```
That is (with marginal error), 
$$
P(Z < yourAnswer) = 0.1357
$$
(b) Fine the Z-score for 63.68 percentile.

```{r}
qnorm(your_answer, mean = 0, sd = 1)
```
That is (with marginal error), 
$$
P(Z < yourAnswer) = 0.6368
$$
### Example 3: arbitrary mean and standard deviation

- As some of you may have guesses both `pnorm` and `qnorm` can be used with arbitrary mean and standard deviation.
- In this case, they internally compute the Z-score from given x and vice versa. 
- For the purpose of Math 2265, students are expected to know how conversion works both ways (x to Z-score and Z-score to x), (the sec 4.1 video (second part) explains how to do this).

### Revisiting the possums example.

The possums example in the video follows the normal distribution with mean 92.6 and sd 3.6. 

When $x=98$, the Z-score is 1.5. With `pnorm`, we computed the percentile (approximately) 0.9331

```{r}
z_98 <- (98-92.6)/3.6
p_98 <- pnorm(z_98, mean = 0, sd = 1)
p_98
```

This process can be combined as 

```{r}
pnorm(98, mean = 92.6, sd = 3.6)
```

When $x=89$, the Z-score is -1, the percentile was (approximately) 0.16. 

```{r}
z_89 <- (89-92.6)/3.6
p_89 <- pnorm(z_89, mean = 0, sd = 1)
p_89
```

### Task 3

Complete the code to compute the percentile for $x=89$.

```{r}
pnorm(your_answer, mean = your_answer, sd = your_answer)
```

---

- We will cover exercises with `qnorm` next time, but it should be enough for up to Q6 in HW4. 
- Also, we will use `xpnorm` to visualize today's code. 

---

### Example 68-95-99.7% Rule

- The percentage within one standard deviation is 68%,
- The percentage within two standard deviations is 95%
- The percentage within one standard deviations is 99.7% 

The percentage within one standard deviation is 68% means 
$$
P( |Z| < 1) = P ( -1 < Z < 1) = 0.68
$$
So, 
$$
P( |Z| < 2) = 0.95
$$
$$
P( |Z| < 3) = 0.997
$$

### Task 4: Compute $P ( |Z| < 2 )$ using `pnorm`. 

```{r}
p_sd2 <- ...pnorm...
p_sd2
```

---

### Task 6. Knit your code and check your outcomes.

You are only allowed to upload pdf or html

### Task 7. Check your answer 

for Task 4

```{r}
if(p_sd2==0.9544997) {
  print("Your answer is correct")
} else {
  print("Check your answer")
}
```


### Share your work and help your group members before uploading your work to Canvas