---
title: "Conditional Probability"
output: html_document
date: "2024-09-24"
author: <your name>
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Math 2265 Section 3.2

You will need to replace `<>` in the source code or answer questions. 

### Load Packages

```{r, echo=FALSE}
library(openintro)
library(ggplot2)
library(tinytex)  # to knit to PDF
set.seed(2265)
```

We will use the `smallpox` data set. From text book p. 99,

> The `smallpox` data set provides a sample of 6,224 individuals from the year 1721 who were exposed to smallpox in Boston. Doctors at the time believed that inoculation, which involves exposing a person to the disease in a controlled form, could reduce the likelihood of death.
> Each case represents one person with two variables: `inoculated` and `result`. The variable `inoculated` takes two levels: `yes` or `no`, indicating whether the person was inoculated or not. The
variable `result` has outcomes `lived` or `died`.

First, use `str` to glance through the data set `smallpox`

```{r, error=TRUE}
str(<>)
```

---

**Question 1.** What are the names of the variables?

**Answer:** 

---

**Question 2.** Use `table` to tally each variable. The syntax is 

For the first variable: 

```{r, error=TRUE}
table(smallpox$<variable>)
```
For the second variable: 

```{r, error=TRUE}
table(smallpox$<variable>)
```

**Question 3.** 

(a) Make a contingency table (two-way table) of the variables `result` and `inoculated` in the data set `smallpox` and save it to a variable `t`.

```{r, error=TRUE}
t <- table(smallpox$<>, smallpox$<>)
t
```

(b) Make a proportional table.

```{r, error=TRUE}
prop.table(t)
```

## Upload your work to Canvas

