---
title: "Conditional Probability"
author: "<your name>"
date: "2024-09-24"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Math 2265 Section 3.2

- Work as a group!
- You will need to replace `<>` in the source code or answer questions. 
- Update your name in L5.

### Load Packages

```{r, echo=FALSE}
library(openintro)
library(ggplot2)
library(tinytex)  # to knit to PDF
set.seed(2265)
```

We will use the `smallpox` data set. From text book p. 99,

> The `smallpox` data set provides a sample of 6,224 individuals from the year 1721 who were exposed to smallpox in Boston. Doctors at the time believed that inoculation, which involves exposing a person to the disease in a controlled form, could reduce the likelihood of death.
> &nbsp
> Each case represents one person with two variables: `inoculated` and `result`. The variable `inoculated` takes two levels: `yes` or `no`, indicating whether the person was inoculated or not. The
variable `result` has outcomes `lived` or `died`.

First, use `str` to glance through the data set `smallpox`

```{r, error=TRUE}
str(<>)
```

---

### Question 1. What are the names of the variables?

**Answer:** 

---

### Question 2. Use `table` to tally each variable. 

For the first variable: 

```{r, error=TRUE}
table(smallpox$<>)
```
For the second variable: 

```{r, error=TRUE}
table(smallpox$<>)
```

### Question 3.

(a) Make a contingency table (two-way table) of the variables `result` and `inoculated` in the data set `smallpox` and save it to a variable `t`.

```{r, error=TRUE}
t <- table(smallpox$<>, smallpox$<>)
t
```

(b) Make a proportional table.

```{r, error=TRUE}
options(digits = 1)
prop.table(<>)
```
## Summary of code

Compared to the Tables 3.15 and 3.16 in the textbook,

- we miss the totals and
- the order of categories are reversed

We can add `addmargins` for the former, but the latter is merely technical (which we won't put much emphasis). The following code will display the table in the same format.

```{r}
smallpox$result <- factor(smallpox$result, levels = c("lived", "died"))
smallpox$inoculated <- factor(smallpox$inoculated, levels = c("yes", "no"))

my_table <- table(smallpox$result, smallpox$inoculated)

print(addmargins(my_table))
options(digits = 2)
print(addmargins(my_table)/sum(my_table))
```

---

### Example/Suggested answer format:

**What is the probability that a randomly selected person lived, given that they were inoculated?**

\[
P(\text{lived} \mid \text{yes}) = \frac{238}{244}
\]

```{r}
238/244
```

### Question 4

---

1. **What is the probability that a randomly selected person was inoculated?**

\[
P(\text{<>}) = \frac{<>}{<>} = 
\]

```{r}

```

---

2. **What is the probability that a person lived and was inoculated?**

\[
P(\text{<>} \cap \text{<>}) = \frac{<>}{<>} = 
\]

```{r}

```

---

3. **What is the probability that a person was inoculated, given that they lived?**

\[
P(\text{<>} \mid \text{<>}) = \frac{<>}{<>} = 
\]

```{r}

```

---

4. **If a randomly selected person died, what is the probability that they were inoculated?**

\[
P(\text{<>} \mid \text{<>}) = \frac{<>}{<>} = 
\]

```{r}

```

---

## Upload your work to Canvas

