---
title: "Conditional Probability"
author: "<your name>"
date: "2024-09-24"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Math 2265 Section 3.2

- Work as a group!
- You will need to replace `<>` in the source code or answer questions
- Update your name in L5
- Make sure your save and `knit` your work (to html or pdf) before submitting it to Canvas

### Question 0. Who are your group members? (List their first names should be sufficient)

**Answer: ** 

  1. <name_1>
  1. <name_2>

### Load Packages

```{r, echo=FALSE}
library(openintro)
library(ggplot2)
library(tinytex)  # to knit to PDF
set.seed(2265)
```

## Example

Review the example and code below to answer the questions listed after the example. There is no question in the example. 

### Expected Value of a Discrete Random Variable

In a game of cards, you win $1 if you draw a heart (excluding the ace of hearts), $5 if you draw an ace (including the ace of hearts), $10 if you draw the king of spades, and nothing for any other card you draw. Write the probability model for your winnings, and calculate your expected winning.

The table below provides the relevant information:

| Event               | Heart (not ace) | Ace      | King of spades | All else |
|---------------------|-----------------|----------|----------------|----------|
| $X$ (reward in $)   | 1               | 5        | 10             | 0        |
| $P(X)$              | 12/52           | 4/52     | 1/52           | 35/52    |

1. Create the variables to store the probability distribution

```{r}
# Define the winnings for X and their respective probabilities P(X)
rewards <- c(1, 5, 10, 0)
probabilities <- c(12/52, 4/52, 1/52, 35/52)
```

2. Use the formula for the expected value:

[
\mathbb{E}(X) = \sum_{i} x_i \cdot P(X = x_i)
]

```{r}
expected_value <- sum(rewards * probabilities)
```

3. Calculate the standard deviation

Hint: First, calculate the variance using the formula:

[
\text{Var}(X) = \sum_{i} (x_i - \mathbb{E}(X))^2 \cdot P(X = x_i)
]

Then, take the square root to find the standard deviation:

[
\sigma_X = \sqrt{\text{Var}(X)}
]

```{r}
# Calculate the variance
variance <- sum((rewards - expected_value)^2 * probabilities)

# Calculate the standard deviation
std_dev <- sqrt(variance)
std_dev
```


---

## Baggage Fees Worksheet

We will work on Problem 2.38 in HW3.

### 2.38 Baggage fees in HW3

> An airline charges the following baggage fees: $30 for the first bag and $30 for the second. Suppose 50% of passengers have no checked luggage, 33% have only one piece of checked luggage, and 17% have two pieces. We assume a negligible portion of people check more than two bags.

Let the random variable $X$ denote the number of pieces of checked luggage. Complete the following table of the probability distribution (we will ignore the `3+` category):

| Number of Luggage | 0   | 1   | 2   |
|-------------------|-----|-----|-----|
| $X$ (Revenue in $)|  0  |  30 |  60 |
| $P(X)$            | 0.5 | 0.33| 0.17|

### Question 1. What is the average baggage-related revenue per passenger?

Complete the code below to store your answer to the variable `expected_value`.

Hint: Use the formula for the expected value:

\[
\mathbb{E}(X) = \sum_{i} x_i \cdot P(X = x_i)
\]


```{r, error=TRUE}
X <- c(0, 30, <>)
prob <- c(0.5, <>, <>)
expected_value <- sum(<>)
expected_value
```

### Question 2. What is the standard deviation of baggage-related revenue?

Complete the code below to store your answer to the variable `expected_value`.

Hint: First, use the formula for variance, and then take the square root to get the standard deviation:

\[
\text{Var}(X) = \sum_{i} (x_i - \mathbb{E}(X))^2 \cdot P(X = x_i)
\]

Then, standard deviation:

\[
\sigma_X = \sqrt{\text{Var}(X)}
\]

```{r, error=TRUE}
my_variance <- sum((<> - <>)^2 * <>)
my_std_dev <- sqrt(my_variance)
std_dev
```

### Question 3. About how much revenue should the airline expect for a flight of 100 passengers?

```{r, error=TRUE}
total_revenue <- expected_value * <>
total_revenue
```

---
### Question 4. Knit your code and check your outcomes.

Did it produce an outcome? 

Yes/No

### Question 5. Check your answer

20.1, 22.47, 2010

---
### Discuss your answers and help your group members before uploading your work to Canvas
