---
title: "Introduction to Normal Distribution"
author: "your name"
date: "2024-10-3"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Math 2265 Sec 4.1 Normal Distribution Part I

-   Work as a group!
-   You will need to replace `"ans"` or `your_answer` in the source code or answer questions
-   Update your name in L3
-   Add your group members name below
-   Make sure your save and `knit` your work (to html or pdf) before submitting it to Canvas

### Question 0. Who are your group members? (List their first names should be sufficient)

**Answer:**

1.  `<name_1>`
2.  `<name_2>`

### Load Packages

```{r, echo=FALSE}
library(openintro)
#library(ggplot2)
if (!tinytex::is_tinytex()) {
  tinytex::install_tinytex()
} # this ensures the necessary package is installed for knit to pdf
library(tinytex)  # to knit to PDF
library(mosaic)
set.seed(2265)
```

------------------------------------------------------------------------

### If you need more time to get used to `Markdown`, use the `Visual` mode.

The icon is located in the upper-left corner next to `source`.

------------------------------------------------------------------------

### Example 1: Plot normal distributions

```{r}
curve(dnorm(x, mean = 0, sd = 1), from = -6, to = 6, 
      main = "Normal Distribution (PDF, mean = 0, sd = 1)",
      xlab = "x", ylab = "Density")
```

### Task 1: Plot some normal distributions with mean and standard deviation of your choice

You may need to change `from` and `to` parameters too.

Your Example 1:

```{r}
curve(dnorm(x, mean = 2, sd = 1.2), from = -6, to = 6)
```

Your Example 2:

```{r}
curve(dnorm(x, mean = -1, sd = 3), from = -6, to = 6)
```

### Example 2: Normal distribution with colors

```{r}
# Plot the normal distribution curve
curve(dnorm(x, mean = 0, sd = 1), from = -4, to = 4, 
      main = "Normal Distribution with Shaded Area (x <= 0)",
      xlab = "x", ylab = "Density")

# Shade the area under the curve for x <= 0
x_vals <- seq(-4, 0, length.out = 100)  # Sequence of x values from -4 to 0
y_vals <- dnorm(x_vals, mean = 0, sd = 1)  # Corresponding y values for the curve
polygon(c(x_vals, 0), c(y_vals, 0), col = "lightblue", border = NA)  # Fill the area
```

This example is excellent but a bit involved. We will use another package `mosaic` for plotting normal distributions.

### Example with the `Mosaic` pacakge

```{r}
xpnorm(0) # if we skip mean and sd, the default values are mean = 0 and sd = 1
```

So the following is the same as above.

```{r}
xpnorm(0, mean = 0, sd = 1)
```

The first argument (0 in the example) is the x-value or Z-score when mean = 0 and sd = 1.

### Z-Score

The possums example in the video follows the normal distribution with mean 92.6 and sd 3.6.

Recall the Z-score formula.

$$
Z = \frac{x-\mu}{\sigma},
$$ where $\mu$ is the mean and $\sigma$ is the standard deviation.

#### Task 2:

Therefore, when $x=98$, the Z-score is

```{r}
(98 -92.6)/3.6
```

Therefore, when $x=89$, the Z-score is

```{r}
(89 -92.6)/3.6
```

### Task 3: Compute the Z-Score and percentile (with table and `pnorm`)

The `pnorm` function will give you the percentile of the Z-score. Here is an example.

```{r}
pnorm(0)
```

Use `pnorm` to compute the percentile for the two Z-score you computed for x = 98 and x = 89.

For x = 98,

```{r}
pnorm(1.5)
```

For x = 89,

```{r}
pnorm(-1)
```

### Task 4: Sketch the normal distribution graphs using the function `xpnorm`

For x = 98,

```{r}
xpnorm(1.5)
```

For x = 89,

```{r}
xpnorm(-1)
```

### Task 5: Try the first HW problem.

Question 1:

**3.1 Area under the curve, Part I**: Find the probability of each of the following, if $Z \sim N(\mu = 0,\sigma = 1)$.

a)  P(Z\<-1.35) =

```{r}
q1_a <- pnorm(-1.35)
q1_a
```

b)  P(Z\>1.48) =

```{r}
q1_b <- 1-pnorm(1.48)
q1_b
```

c)  P(-0.4\< Z\< 1.5) =

```{r}
q1_c <- pnorm(1.5) - pnorm(-0.4)
q1_c 
```

d)  P(\|Z\|\>2) =

```{r}
q1_d <- 2*pnorm(-2)
q1_d
```

------------------------------------------------------------------------

### Task 6. Knit your code and check your outcomes.

You are only allowed to upload pdf or html

### Task 7. Check your answer

```{r}
if(sum((c(q1_a, q1_b, q1_c, q1_d) - c(0.0885,0.0694,0.5886,0.0455))<0.001)==4) {
  print("Your answers are correct")
} else {
  print("Check your answers")
}
```

### Share your work and help your group members before uploading your work to Canvas
